---
title: 'Assignment 3'
author: "Nick McManus"
date: "3 March 2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)

library(tidyverse)
library(purrr)
library(here)
library(kableExtra)
```
# 1) Compute the mean and max power for 4 cars, assuming that average highway speed is 80 kph with a std dev of 10 kph. Do this with both a for loop and `purrr::pmap`.

4 car types:
* car A mass 20000 kg, area 22 m^2
* car B mass 65000 kg, area 30 m^2
* car C mass 38000 kg, area 22 m^2
* car D mass 5000 kg, area 10 m^2

```{r}
### read in auto power and speed conversion functions
source(here("Assignments/R/apg_function.R"))
source(here("Assignments/R/mps_conv_function.R"))

### create dataframe with cars
possible_cars <- data.frame(name = c("A","B","C","D"), 
                            mass = c(10000, 65000, 38000, 5000),
                            area = c(22, 30, 22, 10))

### convert avg speed to mps
base_speed <- mps_conv(80)

possible_cars$power = apg(v=speed_base, 
                          sa = possible_cars$area,
                          m=possible_cars$mass)

# head(possible_cars)
```

First we'll automate this task using a for loop on a sample of 200 log normally distributed speeds. The standard deviation (10kph) is 8% of the mean speed
```{r}
### create sample size
nsample = 200
### convert kph to log mps for average speed
mean_speed = log(base_speed)
### find log normal distribution
speeds = rlnorm(mean=mean_speed, sd=0.08*mean_speed, nsample)

### explore speed data
# summary(speeds)
# plot(density(speeds), ylab="Distribution of Speeds in (m/s)")

### create empty dataframe to store for loop results
### 4 columns for each car and 200 rows for each observed speed
powerall = as.data.frame(matrix(nrow = length(speeds),
                                ncol = nrow(possible_cars)))

### now run for loop using data from `possible_cars`
for (i in 1:length(powerall)) {
  powerall[,i] = apg(m = possible_cars$mass[i],
                     v = speeds,
                     sa = possible_cars$area[i]) 
}

### add column names
colnames(powerall) = possible_cars$name

### turn columns in to rows for plotting and finding means
powerallr = powerall %>% 
  pivot_longer(cols = everything(), 
               names_to = "car", 
               values_to = "power")

### Find mean powers and return in nice table
means_floop <- powerallr %>% 
  group_by(car) %>% 
  summarize(mean = mean(power),
            max = max(power))

means_floop %>% 
  kable(caption = "**Table 1.** Mean and max powers for cars A-D using for-loop method",
        col.names = c("Car", "Mean power", "Max power")) %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = "hover",
                position = "left")
```


Next, we'll find the mean and max on the same sample size of cars and speeds using `purrr::pmap`
```{r}
### use same speed data as for loop

### run pmap
powerall2 = pmap(list(sa = possible_cars$area,
                      m = possible_cars$mass), 
                 apg, v = speeds)

### convert pmap results into dataframe
powerall2 = as.data.frame(powerall2, 
                          col.names = possible_cars$name)

### turn columns in to rows for plotting and finding means
powerallr2 = powerall2 %>% 
  pivot_longer(cols = everything(), 
               names_to = "car", 
               values_to = "power")

### Find mean powers and return in nice table
means_pmap <- powerallr2 %>% 
  group_by(car) %>% 
  summarize(mean = mean(power),
            max = max(power))

means_pmap %>% 
  kable(caption = "**Table 2.** Mean and max powers for cars A-D using pmap method",
        col.names = c("Car", "Mean power", "Max power")) %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = "hover",
                position = "left")
```


# 2) Create two boxplots (one for each method) that show the range of power consumption across sampled speeds by car type.

```{r, fig.height=8, fig.width = 5}

forLoop_plot <- ggplot(data = powerallr, 
                       aes(x = car, y = power)) +
  geom_jitter(aes(color = car), width = .13, alpha = 0.7, size = 1.5) +
  scale_color_manual(values = c("sienna1", "mediumorchid1", "aquamarine2", "firebrick1")) +
  geom_boxplot(fill = NA, outlier.color = NA, width = 0.6, lwd = 0.8, 
               color = c("sienna4", "mediumorchid4", "aquamarine4", "firebrick4")) +
  labs(x = "Car",
       y = "Power (Watts)",
       title = "For loop") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    axis.title.x = element_text(vjust = 5, face = 'bold'),
    axis.title.y = element_text(face = 'bold'),
    axis.text.x = element_text(vjust = 7, face = 'bold', size = 9),
    axis.text.y = element_text(size = 8),
    legend.position = 'none')



pmap_plot <- ggplot(data = powerallr2, 
                    aes(x = car, y = power)) +
  geom_jitter(aes(color = car), width = .13, alpha = 0.7, size = 1.5) +
  scale_color_manual(values = c("sienna1", "mediumorchid1", "aquamarine2", "firebrick1")) +
  geom_boxplot(fill = NA, outlier.color = NA, width = 0.6, lwd = 0.8, 
               color = c("sienna4", "mediumorchid4", "aquamarine4", "firebrick4")) + 
  labs(x = "Car",
       y = "Power (Watts)",
       title = "pmap") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    axis.title.x = element_text(vjust = 5, face = 'bold'),
    axis.title.y = element_text(face = 'bold'),
    axis.text.x = element_text(vjust = 7, face = 'bold', size = 9),
    axis.text.y = element_text(size = 8),
    legend.position = 'none')

ggarrange(forLoop_plot, pmap_plot, 
          labels = c("A", "B"), 
          hjust = -2,
          nrow=2)
```

