---
title: 'Assignment 3'
author: "Nick McManus"
date: "3 March 2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)

library(tidyverse)
library(purrr)
library(here)
library(kableExtra)
```



4 car types:

* car A mass 20000 kg, area 22 m^2
* car B mass 65000 kg, area 30 m^2
* car C mass 38000 kg, area 22 m^2
* car D mass 5000 kg, area 10 m^2


1) Compute the mean and max power for the 4 cars, assuming that average highway speed is 80 km/hr with a std dev of 10 km/hr
- Do this with both a for loop and purrr::pmap


```{r}
### read in auto power and speed conversion functions
source(here("Assignments/R/apg_function.R"))
source(here("Assignments/R/mps_conv_function.R"))

### create dataframe with cars
possible_cars <- data.frame(name = c("A","B","C","D"), 
                            mass = c(10000, 65000, 38000, 5000),
                            area = c(22, 30, 22, 10))

### convert avg speed to mps
base_speed <- mps_conv(80)

possible_cars$power = apg(v=speed_base, 
                          sa = possible_cars$area,
                          m=possible_cars$mass)

# head(possible_cars)
```

First we'll automate this task using a for loop on a sample of 200 log normally distributed speeds. The standard deviation (10kph) is 8% of the mean speed
```{r}
### create sample size
nsample = 200
### convert kph to log mps for average speed
mean_speed = log(base_speed)
### find log normal distribution
speeds = rlnorm(mean=mean_speed, sd=0.08*mean_speed, nsample)

### explore speed data
# summary(speeds)
# plot(density(speeds), ylab="Distribution of Speeds in (m/s)")

### create empty dataframe to store for loop results
### 4 columns for each car and 200 rows for each observed speed
powerall = as.data.frame(matrix(nrow = length(speeds),
                                ncol = nrow(possible_cars)))

### now run for loop using data from `possible_cars`
for (i in 1:length(powerall)) {
  powerall[,i] = apg(m = possible_cars$mass[i],
                     v = speeds,
                     sa = possible_cars$area[i]) 
}

### add column names
colnames(powerall) = possible_cars$name

### turn columns in to rows for plotting and finding means
powerallr = powerall %>% 
  pivot_longer(cols = everything(), 
               names_to = "car", 
               values_to = "power")

### Find mean powers and return in nice table
means_floop <- powerallr %>% 
  group_by(car) %>% 
  summarize(mean = mean(power),
            max = max(power))

means_floop %>% 
  kable(caption = "**Table 1.** Mean and max powers for cars A-D using for-loop method",
        col.names = c("Car", "Mean power", "Max power")) %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = "hover",
                position = "left")
```


Next, we'll find the mean and max on the same sample size of cars and speeds using `purrr::pmap`
```{r}

```




```{r}
# quick visualization
forLoop_plot = ggplot(powerallr, aes(car,power, fill=car))+geom_boxplot()+ggtitle("For Loop")
forLoop_plot

ggarrange(method1_plot, method2_plot, method3_plot, nrow=3)
```

